---
title: "p8105_hw2_rq2166"
author: "Ruoyuan Qian"
date: "2019/9/28"
output: github_document
---
```{r setup, message=FALSE}
library(tidyverse)
library(readxl)
library(dplyr)
```

# Problem 1

Read and clean the `Mr. Trash Wheel sheet`
```{r read and clean the Mr. Trash Wheel sheet}

## read the specific dataset with reasonable column names
## omit rows not include specific values of dumpster
## round sports ball to the nearest integer and convert to integer value
Trash_Wheel = 
   read_excel( "./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",
               sheet = "Mr. Trash Wheel",
               range = "A2:N338")%>%
   janitor::clean_names()%>%
   drop_na( dumpster )%>%
   mutate( sports_balls = round( sports_balls,0 ),
           sports_balls = as.integer( sports_balls ))
```

Read and clean `precipitation data` for 2017 and 2018
```{r}
precipitation_2018 = 
  read_excel( "./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",
             sheet = "2018 Precipitation",
             skip  = 1 )%>%
  janitor::clean_names()%>%
  drop_na()%>%
  mutate( year = 2018 )

precipitation_2017 = 
  read_excel( "./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",
             sheet = "2017 Precipitation",
             skip  = 1 )%>%
  janitor::clean_names()%>%
  drop_na()%>%
  mutate( year = 2017 )

## combine 2 datasets
precipitation = bind_rows( precipitation_2018,
                           precipitation_2017 )%>%
                mutate( month = as.integer(month),
                        month = month.name[month]) %>% 
                select(year,month,everything())
```
**`Description:`**

`Mr. Trash Wheel`

The number of observation in `Trash_Wheel` is ``r Trash_Wheel%>%pull(dumpster)%>%length()``. 

The key variables in `Mr. Trash Wheel` spreadsheet include information on the dumpter number(`dumpster`), date of collection(`year`, `month`, `date`), amount of total litter(`weight_tons`, `volume_cubic_yards`) and litter type(`Plastic Bottles`, `Glass Bottles`, `Grocery Bags`, `Chip Bags`, `Sports Balls`). Additionally, it contians Homes Powered(`Homes Powered*`) - Each ton of trash equates to on average 500 kilowatts of electricity.


`precipitation`

The number of observation in `precipitation` is ``r precipitation%>%pull(year)%>%length()``. The key variable in  `precipitation` is the amount of precipitation(`total`) in each month(`month`) in different years(`year`). 


**`Total and Median`**

The `total` precipitation in 2018 is ``r precipitation %>%filter(year=="2018")%>%pull(total)%>%sum``, the `median` number of `sports balls` in a dumpster in 2017 is ``r Trash_Wheel %>%filter(year==2017) %>%pull(sports_balls)%>%median``.



# Problem 2

Clean the data in `pols-month.csv`
```{r ,message=FALSE}
pols_month = 
  read_csv ( "./data/fivethirtyeight_datasets/pols-month.csv" )%>%
  separate ( mon, c("year","month","day"), "-" )%>%
  pivot_longer(
    c( prez_gop,prez_dem ),
    names_to = "president",
    names_prefix = "prez_" )%>%
  filter ( value != 0) %>%
  select ( -day,-value)%>%
  mutate (
    month = as.integer(month),
    month = month.name[month])

```


Clean the data in `snp.csv`
```{r ,message=FALSE}
snp = 
  read_csv( "./data/fivethirtyeight_datasets/snp.csv" )%>%
  separate( date, c("month","day","year"), "/" )%>%
  select( -day ) %>%
  mutate(
    month = as.integer(month),
    month = month.name[month]) %>% 
  select ( year,month,everything()) %>%
  arrange( year,month)
  

```


Tidy the `unemployment.csv`
```{r ,message=FALSE}
unemployment = 
  read_csv( "./data/fivethirtyeight_datasets/unemployment.csv" )%>%
  pivot_longer(
    Jan:Dec,
    names_to  = "month",
    values_to = "unemployment" )%>%
  rename ( year = Year)%>%
  arrange( year,month )%>%
  mutate ( year  = as.character(year),
          month  = match(month,month.abb),
          month  = month.name[month] )
```


Merge `snp` into `pols` , and merging `unemployment` into the `result`
```{r}
  tem_result = left_join( pols_month , snp,
                        by=c("year","month"))

  result     = left_join( tem_result , unemployment,
                        by=c("year","month"))
```

**Summary of `pols_month`:** 

The variable number is ``r pols_month %>%length()``, the number of observation ``r pols_month %>%pull(year)%>%length()``, the `range of years` is from ``r range(pols_month%>%pull(year))[1]`` to ``r range(pols_month%>%pull(year))[2]``. 
###The key variables are `year` and `month`.

**Summary of `snp`:** 

The variable number in `snp` is ``r snp%>%length()``. The number of observation in `snp` is ``r snp%>%pull(year)%>%length()``. The  `range of years` in `snp` is  from ``r range(snp%>%pull(year))[1]`` to ``r range(snp%>%pull(year))[2]``.

**Summary of `unemployment`:** 

The variable number in `unemployment` is ``r unemployment%>%length()``. The number of observation in `unemployment` is ``r unemployment%>%pull(year)%>%length()``.The `range of years` in `unemployment` is from ``r range(unemployment%>%pull(year))[1]`` to `range(unemployment%>%pull(year))[2]`. 

The names of **key variables** in `pols-month`, `snp` and `unemployment` are `year` and `month`.


**Summary of `result`:** 

In result, the number of observations and variables is ``r result%>%dim()``, respectively. The range of years is from``r range(result%>%pull(year))[1]`` to ``r range(result%>%pull(year))[2]``, the key variables are `year` and `month`.


# problem 3

Tidy data `Popular_Baby_Names.csv`
```{r ,message=FALSE}
## change ethnicity and childs_first_name into consistent formats respectively
## remove duplicate rows
Popular_Baby_Names = 
  read_csv("./data/Popular_Baby_Names.csv")%>%
  janitor::clean_names()%>% 
  mutate(
     ethnicity = 
       recode(ethnicity, 
               "ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER",
               "BLACK NON HISP" = "BLACK NON HISPANIC",
               "WHITE NON HISP" = "WHITE NON HISPANIC"),
     childs_first_name = str_to_title ( childs_first_name )) %>% 
  distinct()
```

Table of popularity of the name “Olivia” 
```{r ,message=FALSE}
popularity_of_olivia =
  Popular_Baby_Names %>% 
  filter(childs_first_name == "Olivia") %>% 
  select(-count) %>% 
  pivot_wider(names_from  = year_of_birth, 
              values_from = rank) %>%
  select(-gender,-childs_first_name) %>% 
  knitr::kable(
               caption = "Table 1  The rank in popularity of the name “Olivia” as a female baby name over time ",
               format = "html")
  
  popularity_of_olivia
```


Table of the most popular name among male children over time
```{r ,message=FALSE}

popular_name_of_male =
  Popular_Baby_Names %>% 
  filter(rank == 1 & gender == "MALE") %>% 
  select(-count) %>% 
  pivot_wider(
    names_from  = year_of_birth, 
    values_from = childs_first_name) %>% 
  select(-gender,-rank) %>% 
  knitr::kable(
               caption = "Table 2  The most popular name among male children over time ",
               format = "html")
popular_name_of_male
```

```{r}
Popular_Baby_Names %>%  
  filter(year_of_birth == 2016 & 
         ethnicity == "WHITE NON HISPANIC" &
         gender == "MALE") %>% 
  ggplot( aes (x = rank , y = count ,size=count) ) +
  geom_point(color = 4,alpha = .6) +
  labs(
       x    = "Rank in popularity",
       y    = "Number of childern with a name",
       title = "Plot 1  The number with a name against the rank in popularity", 
       subtitle = "Male, white non-hispanic children born in 2016")
  
```



